\newcommand{\motivation}[0]{

{\Huge \textbf{Chapter 1}}\\
	
\section{Introduction}
%	\Large\textbf{\\}
%	{\Huge \textbf{Introduction}}\\
	\subsection{Justification of intererest and relevance}
	 In this work we want to partially answer question of how density-based algorithms can be applying to figure out how visible matter(1) is shaped arraound our position in the Universe.
	 
	 Now a days we all are furtunated to have public Surveys such as SDSS (Sloan Digital Sky Survey) and 2DFGRS ( Two-degree Field Galaxy Redshidft survey) which producing data accesible for everyone. These data contain among others thigs, rows representing galaxies with their redshifts. By applying Universe-models like Labmda-CDM to these data it is possible to estimate the distance of such objects to Earth. 
	 
	 And there is where non-supervised methods can be suitable to make analysis, specially density-based algorithms 
	  This question can be adressed througth density-based algorithms. 
	 	 
	 When we look at the universe arround us, we see some kind of structure: galaxies groupped into clusters and clusters in turn groupped into superclusters and this lasts into structures still bigger. 
	 
	 Density-based unsupervised methods should be able to find the same structure: galaxy clusters integrated on superclusters.
	 
	 \subsection{Personal motivation}
	 
	 As a Master Degree in Data Science, AstroIT developer and 
	
	 Since the beginning of time, the man has always looked at the nigth sky and felt the feeling of the inmense and profound dark of the Universe. This lead us o the question of our postion within the Universe.
	 Now a days, sets of data are been collected directly from satellites and published, making them accessible to everyone via internet. Modern Machine Learning techniques can be apply to this data sets to yield interestring insights. That is main aim of this paper.
}
\input{front.tex}