\documentclass[12pt,a4paper,twoside]{book}  % add one blank page between sections
%\documentclass[12pt,a4paper,oneside]{book} % no more spaces between sections
\usepackage{graphicx}
\usepackage{setspace} %double spacing for text, single for captions, footnotes, etc.
%\usepackage{hypernat} %substitute for cite that allows hyperlinks
\usepackage{natbib} % substitute for 'hypernat' that works on Windows.
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{hhline} % extended styles for tables
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{acronym}
\usepackage{hyperref}
\usepackage{amsmath,amsmath,amssymb}
\usepackage{fancyhdr}
\usepackage{epsfig, amsmath}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage[linesnumbered, ruled]{algorithm2e}
\usepackage{epigraph} 

% general settings
\hypersetup{
linktocpage=true,
colorlinks=true,
linkcolor=blue,
citecolor=blue,
}
\definecolor{Hgray}{gray}{0.6}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\setlength{\topmargin}{0cm}
\setlength{\textheight}{23cm}
\setlength{\textwidth}{17cm}
\setlength{\oddsidemargin}{0cm}
\setlength{\evensidemargin}{0cm}
\setlength{\headheight}{1cm}

% indicates that 'sub-sub-sections' are numbered and appear in the index
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{2}

% settings for code
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

%%%%%%%%%%%%
% DOCUMENT %
%%%%%%%%%%%%
\begin{document}

\setcounter{section}{0} % Resets the section counter to 0 at the document's start
\renewcommand{\thesection}{\arabic{section}} % Changes the section numbering scheme

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Definning new variables			%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\titulo}{Applying Density-Based Algorithms to Galaxy Cluster Catalogs}
\newcommand{\subtitulo}{Unveiling Galaxy Structure with Unsupervised Clustering and 2PFC}
\newcommand{\autor}{Carlos Toro Peñas}
\newcommand{\supervisor}{Laura Ruiz Dern}
\newcommand{\area}{4: Data Science}
\newcommand{\profesor}{David Masip Rodo}
\newcommand{\keywords}{clustering, galaxy clusters, cosmology}
\newcommand{\keywordses}{clustering, cúmulos de galaxias, cosmología}

% cover page
\input{sections/0_title.tex}
\newpage

% abstract
\input{sections/0_abstract.tex}
\newpage

\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{ \markboth{#1}{}}
\renewcommand{\sectionmark}[1]{\markright{ \thesection.\ #1}}

\lhead[\fancyplain{}{\bfseries\thepage}]{\fancyplain{}{\bfseries\rightmark}}
\rhead[\fancyplain{}{\bfseries\leftmark}]{\fancyplain{}{\bfseries\thepage}}
\cfoot{}

% table of contents
%\cleardoublepage
%\clearpage
\phantomsection
\addcontentsline{toc}{chapter}{Table of Contents}
\tableofcontents
% list of figures
%\cleardoublepage
%\clearpage
\phantomsection
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

% list of tables
%\cleardoublepage
%\phantomsection
%\addcontentsline{toc}{chapter}{List of Tables}
%\listoftables

\thispagestyle{empty}

\pagenumbering{arabic}

%\pagestyle{fancy}
%\renewcommand{\chaptermark}[1]{ \markboth{#1}{}}
%\renewcommand{\sectionmark}[1]{\markright{ \thesection.\ #1}}
%\lhead[\fancyplain{}{\bfseries\thepage}]{\fancyplain{}{\bfseries\rightmark}}
%\rhead[\fancyplain{}{\bfseries\leftmark}]{\fancyplain{}{\bfseries\thepage}}
%\cfoot{}


% chapters of the document

%\cleardoublepage
\clearpage
% Introduction
\input{sections/1_introduction.tex}
\newpage

\clearpage

% Introduction
% state of the art chapter
\input{sections/4_state_of_art.tex}

\newpage
% Other subsections from template are still possible to be included
%\input{/sections/3_other_subsections.tex}
\input{sections/5_implementation.tex}

\input{sections/6_conclusions.tex}
%\section{Results}

%-

%\section{Conclusions and future work}

%-

\phantomsection
\pagestyle{fancy}
\chapter{Glossary}
%\section{Glossary}
%-
\begin{description}
    \item[BAO (Baryon Acoustic Oscillations):] Periodic fluctuations in the density of the visible baryonic matter of the universe, acting as a "standard ruler" for cosmological distances.
	
	\item[CasJobs SQL interface:] Online batch-processing system that gives users access to the multi-terabyte SDSS catalog. Unlike simple web forms that allow for small data downloads, CasJobs is designed for Large-scale Data Mining.
	
	\item[$\Lambda$CDM Model:] The current standard model of cosmology, representing a Universe dominated by Dark Energy ($\Lambda$) and Cold Dark Matter (CDM).
	
	\item[Dark Matter Halo:] A quasi-equilibrium state of dark matter particles. The gravitational "wells" where galaxies and galaxy clusters form and reside. See the introduction section in \cite{Hai-Xia-Ma:2025}.
	
	\item[Kaiser effect:] distorsion in apparent clustering of galaxies that appeaers caused by motions of galaxies as they fall into large structures, causing a "squishing" or flattening along the line of sight. See \cite{Shi:2016} for more details.
	
	\item[$k$-d Tree (k-dimensional Tree):] A space-partitioning data structure used to organize points in a $k$-dimensional space. It allows for high-speed "nearest neighbor" searches, essential for large datasets.
	
	\item[Poisson Distribution:] A random spatial distribution where points are placed independently. This serves as the "null hypothesis" against which the 2dFGRS clustering is measured.
	
	\item[Redshift:] Increase in the wavelength of radiation - tipically ligth-. The redshift takes place for several reasons, one of then is when the source of ligth is further away, for example in an expanding Universe, then they speak about cosmic-redshift.
	
	\item[Redshift Space Distortions (RSD):] An effect where the observed distance of a galaxy is distorted by its peculiar velocity (movement due to gravity), causing clusters to look elongated ("Fingers of God") \cite{Shi:2016}.

	\item[Z-score Normalization (Standardization):] A preprocessing step that transforms data to have a mean of 0 and a standard deviation of 1, preventing one feature (like survey depth) from dominating the distance calculation.
	
\end{description}

\phantomsection
\pagestyle{fancy}
% \chapter{Bibliography}

\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{plain}
\bibliography{refs}

%Appendices section commented for now
% \section{Appendices}

\end{document}