\newcommand{\motivation}[0]{

{\Huge \textbf{Chapter 1}}\\
	
\section{Introduction}
%	\Large\textbf{\\}
%	{\Huge \textbf{Introduction}}\\
	\subsection{Justification of intererest and relevance}

	 In this work we want to get an approximation of how density-based algorithms can be applied to detect clusters of galaxies in order to determine the shape of visible matter(1) arround our position in the Universe.
	 
	 Today we are all fortunate to have public surveys such as the Sloan Digital Sky Survey (SDSS) and the Two-degree Field Galaxy Redshift (2DFGRS) that produce data accessible to everyone through internet. These data contain among others things, thousands of rows representing galaxies with their redshifts. By applying Universe-models like Labmda-CDM to these data it is possible to estimate the distance of such objects to Earth. 
	 
	 How dense the matter is? how galaxies are distributed along the Universe? Is it possible to answer these questions using density-based algorithms?
	 	 
	 Density-based unsupervised methods should be able to find the same structure: galaxy clusters integrated on superclusters.
	 
	 \subsection{Personal motivation}
	 
	 Since i was a child I have been deeply interested in the world of galaxies, later, as a master degree in Data Science student I realized that Maching Learing methods can be aplied to a wide variety of fields. In particular, density-based algorithms can guess how clusters and superclusters of galaxies are arranged across the Universe and show us, at least at a quantitative way, how the matter is disposed arround us.
	
	\subsection{Objectives}
	  \begin{itemize}
	   \item An alternative way to guess how matter is distributed in a big scale arround our galaxy the Via Lactea.
	   \item Measurements on galaxy clusters detection througth different density-based algorithms.
	   \item Comparison of how well different density-based algorithms perform clusters dertection.
	  \end{itemize}
	  
	 \subsection{Methodology and project deployment}
	For the present job, I use a set of packages based on R programming. Algorithms I want to compare are:
		OPTICS
		DBSCAN
		HDBSCAN
	
}
\input{front.tex}